---
title: "R Notebook"
output: html_notebook
---
```{r}
# Loading the requireed libraries
library(mlogit)
```

# Choice-Based Conjoint Anlysis

In the following R Markdown, a dataset containing the answers of a choice-based conjoint survey in order to understand the prefereces of customers relatively to the Trading Apps. 

## Data Exploration. 
Let's start by loading the dataset 
```{r}
df <- read.csv('C:/Users/alber/Desktop/UniTn/Data Science/Third Semester/Laboraotry of Business and Customer analytics/Project_Real/survey_results.csv')
```
Now that the dataset is loaded, it is possible to do a bit of data exploration. 
```{r}
dim(df)
```
The dataset  contains 863 observations and 11 columns.
```{r}
names(df)
```
The above are the names of the columns of the dataset. 
We can drop both `X` and `index`. 
```{r}
df <-  df[,-c(1)]
```
Now that unuseful columns have been removed, it is possible to a closer look to the data. 
```{r}
summary(df)
```

Let's compute the descriptive statistics 
```{r}
xtabs(choice ~ Platform, data=df)
xtabs(choice ~ Deposit, data=df)
xtabs(choice ~ Fees, data=df)
xtabs(choice ~ Financial_Instrument, data=df)
xtabs(choice ~ Leverage, data=df)
xtabs(choice ~ Social_Trading, data=df)
xtabs(choice ~ position, data=df)

```
Now, it is possible to recode some variables
```{r}
df$Platform <- as.factor(df$Platform)
df$Deposit <-  as.factor(df$Deposit)
df$Fees <-  as.factor(df$Fees)
df$Financial_Instrument <-  as.factor(df$Financial_Instrument)
df$Leverage <-  as.factor(df$Leverage)
df$Social_Trading <-  as.factor(df$Social_Trading)
```


It is now time to fit a choice model. In order to do this, it is necessary to transform the dataset. 
```{r}
df.mlogit <- dfidx(df, idx = list(c("resp.id","ques"), "position"))

```







